---
title: "Topic 5 Problem Set"
output: html_document
author: Xue (Tracy) Feng
---

## Preparation

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
library(tidyverse)
library(tidytext)
library("sentimentr")
library(wordcloud)
```

The `sentimentr` package includes several text data sets.  For this exercise use the "hotel_reviews" data.  This data set contains 5,000 hotel reviews scraped from TripAdvisor.  It also contains a sentiment score for each review, a score produced via automated sentiment analysis.

Access this data by installing the `sentimentr` package, then loading it via `library(sentiment)` and (for convenience) creating a named data object (e.g., `rev = sentimentr::hotel_reviews`).

```{r echo = FALSE}
rev = sentimentr::hotel_reviews
```


#### (a) Calculate the following:
#### The number of reviews that include the terms "price" or "value"
#### The number of reviews that include the terms "clean" or "neat" or "dirty" or "gross"

```{r echo = TRUE}

# make sure to include all lower and upper cases words
# rev$textLower <- tolower(rev$text)

sum(str_detect(rev$text, "price|value"))

sum(str_detect(rev$text, "clean|neat|dirty|gross"))

```


#### (b) Some of the reviews include superfluous text in the form of "showReview(1132073, 'full);", (the specific number is different across reviews).  Remove this phrase from all the reviews.

```{r echo = TRUE}

# lines number that contain the string we want to remove:
# str_which(rev$text, pattern = "showReview\\(\\d+\\,.+;")

# before removal
rev$text[4031]

# after removal
rev$textRm <- str_remove_all(rev$text, "showReview\\(\\d+\\,.+;")
rev$textRm[4031]

```


#### (c) The reviews have been analyzed for sentiment (the "sentiment" variable in the data set). Show a table that shows distribution of sentiment variable. Look at example text for the most negative and second most negative categories and discuss implications.


```{r}

rev %>%
  group_by(sentiment) %>%
  summarise(Count = n(),
            `Percent Sentiment` = n() / nrow(rev) * 100)

# examples for the most negative category (-2.5)
rev %>%
  filter(sentiment == -2.5) %>%
  select(text) %>%
  head(5)

# examples for the second most negative category (-1.5)
rev %>%
  filter(sentiment == -1.5) %>%
  select(text) %>%
  head(5)

# implications
# The most negative sentiment reviews are not even actual reviews of the hotel. Sometimes its just a person with a bad taste in their mouth leaving a negative review and not justifying it with any facts. These are not reliable reviews.
# The 2nd most negative reviews appear to be more reliable complaints from the hotel guests.

```


#### (d) Calculate the proportion of reviews with sentiment = -1.5 that referenced our cleanliness terms above. 

```{r}

mean(str_detect(subset(rev, sentiment == -1.5)$text, "clean|neat|dirty|gross")) * 100

```


#### (e) Create a word count for the negative sentiment reviews (those coded as sentiment equal to -1.5).  Show the most common words.
```{r message = FALSE}

rev %>%
  filter(sentiment == -1.5) %>%
  unnest_tokens(word, text) %>%
  count(word, sort = T) %>%
  head()

```

#### (f) Show a word cloud for the negative sentiment reviews (those coded as sentiment equal to -1.5) with stopwords removed.
```{r message = FALSE}

rev %>%
  filter(sentiment == -1.5) %>%
  unnest_tokens(word, text) %>%
  anti_join(tidytext::stop_words, by = "word") %>%
  filter(!str_detect(word, "\\d")) %>%
  count(word) %>%
  with(
    wordcloud::wordcloud(
      word,
      n,
      max.words = 35,
      random.order = FALSE,
      colors = brewer.pal(8, "Dark2")
    )
  )

```

#### (g) Show a word cloud for the positive sentiment reviews (those coded as sentiment equal to 2.5) with stopwords removed.

#### Since _hotel_ is not particularly informative, delete it as a stop word.  (Use `anti_join(rbind("hotel", tidytext::stop_words))` at the stop words stage.)

```{r message = FALSE}

rev %>%
  filter(sentiment == 2.5) %>%
  unnest_tokens(word, text) %>%
  anti_join(rbind("hotel", tidytext::stop_words), by = "word") %>%
  filter(!str_detect(word, "\\d")) %>%
  count(word) %>%
  with(
    wordcloud::wordcloud(
      word,
      n,
      max.words = 35,
      random.order = FALSE,
      colors = brewer.pal(8, "Dark2")
    )
  )

```

#### (h) We are interested in finding words that are most distinctive in each sentiment group.

#### We're going to need to do some more cleaning.  First, we need to filter out the reviews coded as -2.5 sentiment as these are not real reviews.

#### Second, we're going to count the words, grouped by sentiment.  To do this, use `count(sentiment, word, sort = TRUE)` which will count the word usage broken into sentiment categories.  Display the top 25 terms sorted by tf-idf in descending order.


```{r message = FALSE}

rev %>%
  filter(sentiment != -2.5) %>%
  unnest_tokens(word, text) %>%
  # It's unclear in the question whether we should include the word "hotel" or not
  # anti_join(rbind("hotel", tidytext::stop_words, by = "word")) %>%
  anti_join(rbind(tidytext::stop_words, by = "word")) %>%
  count(sentiment, word, sort = TRUE) %>%
  group_by(sentiment) %>%
  mutate(total = sum(n)) %>%
  bind_tf_idf(word, sentiment, n) %>%
  arrange(-tf_idf) %>%
  head(25)

```

