---
title: "Topic 2 Lab"
output: html_document
author: Insert Your Name Here
---

## Preparation

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
# Load packages used in this session of R
library(tidyverse)
library(knitr)
library(dplyr)

# As needed, set path to folder where data is located.
opts_knit$set(root.dir = "C:/Users/baileyma/Dropbox/Teaching_drop/OPIM606")
```


#### 1) Load the World Values Survey data from "Data/WV_small.csv" and the CountryCode data from "Data/Country codes for WVS wave 6.csv". Create an age variable.

The codebook for the World Values Survey data is available at 
http://www.worldvaluessurvey.org/WVSDocumentationWV6.jsp

The variables in WV_small include

- Satisfaction with your life from 1 (completely dissatisfied) to 10 (completely satisfied) (*V23*)
- Income: a country-specific scale ranging from 1 (lowest income category) to 10 (highest income category) (*V239* in the dataset)
- Education: a scale ranging from 1 (no formal education) to 9 (a degree from a university) (*V248* in the dataset)
- Country: based on *V2* in the dataset.  See “Country codes for WVS wave 6.csv”
- Conservatism: self-identified political conservatism from 1 (most liberal) to 10 (most conservative) (*V95* in the data set)
- Male: a dummy variable indicating male (*V240*)
- Religious: Indicating how often the individual attends religious services ranging from 1 (almost never) to 7 (more than once a week) (this is a re-coding of *V145* that had the polarity reversed, but is harder to interpret)
- Marital: Marital status 1 = Married 2 = Living together as married 3 = Divorced 4 = Separated 5 = Widowed 6 = Single
- Survey year: year survey taken
- Birth year: respondents's year of birth


```{r tidy = FALSE}

```

#### **After** creating the WV and CountryCode data objects, implement the following to create a list of countries in the WV data.  (We'll cover what this code means later in data wrangling.)  Look at this data object to understand it.  It will be useful in the last question below.

```{r }
# Load country code data
CountryCode = read.csv("Data/Country codes for WVS wave 6.csv")
WV = read.csv("Data/WV_small.csv")

wvCountries =  CountryCode %>% 
  filter(V2>0) %>%
  filter(V2 %in% unique(WV$Country)) %>%
  filter(V2 != 332)
```

#### 2) Create a list that contains a country name, the average satisfaction in that country and the observations for Mexico.

```{r tidy = FALSE}
```

#### 3) Estimate a model predicting *DemocracyBad* (whether someone thinks democracy is bad) and save it as OLS.1.  Briefly explain the structure of this object. What is OLS.1[[1]]?

```{r tidy = FALSE}
```
#### 4) Loop thru all countries in the World Values survey to assess the bivariate relationship between opposition to democracy and age. Save coefficient and t-statistic for each country.


```{r tidy = FALSE}
```


#### 5) Create a function that prints a table with response percent for each answer for *DemocracyBad* question for any country

Recall that the answer categories are:

1. Very good
2. Fairly good
3. Fairly bad
4. Very bad  


```{r}
```

#### 6) Create a list with regression results for an OLS model explaining *DemocracyBad* as a function of age, income, education and male.  Be able to print results by country name.

```{r}
```
