---
title: "Topic 2 Lab"
output: html_document
author: Xue(Tracy) Feng
---

## Preparation

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
# Load packages used in this session of R
library(tidyverse)
library(knitr)
library(dplyr)
library(hash)
library(plyr)

# As needed, set path to folder where data is located.
opts_knit$set(root.dir = "C:/Users/Xue Feng/Desktop/OPIM606/Lab2")
```


#### 1) Load the World Values Survey data from "Data/WV_small.csv" and the CountryCode data from "Data/Country codes for WVS wave 6.csv". Create an age variable.

The codebook for the World Values Survey data is available at 
http://www.worldvaluessurvey.org/WVSDocumentationWV6.jsp

The variables in WV_small include

- Satisfaction with your life from 1 (completely dissatisfied) to 10 (completely satisfied) (*V23*)
- Income: a country-specific scale ranging from 1 (lowest income category) to 10 (highest income category) (*V239* in the dataset)
- Education: a scale ranging from 1 (no formal education) to 9 (a degree from a university) (*V248* in the dataset)
- Country: based on *V2* in the dataset.  See “Country codes for WVS wave 6.csv”
- Conservatism: self-identified political conservatism from 1 (most liberal) to 10 (most conservative) (*V95* in the data set)
- Male: a dummy variable indicating male (*V240*)
- Religious: Indicating how often the individual attends religious services ranging from 1 (almost never) to 7 (more than once a week) (this is a re-coding of *V145* that had the polarity reversed, but is harder to interpret)
- Marital: Marital status 1 = Married 2 = Living together as married 3 = Divorced 4 = Separated 5 = Widowed 6 = Single
- Survey year: year survey taken
- Birth year: respondents's year of birth

#### **After** creating the WV and CountryCode data objects, implement the following to create a list of countries in the WV data.  (We'll cover what this code means later in data wrangling.)  Look at this data object to understand it.  It will be useful in the last question below.

```{r }
# Load country code data
countryCode = read_csv("Country codes for WVS wave 6.csv")
wv = read_csv("WV_small.csv")

# age variable
wv <- wv %>%
  mutate(age = SurveyYear - BirthYear)

wvCountries =  countryCode %>% 
  filter(V2>0) %>%
  filter(V2 %in% unique(wv$Country)) %>%
  filter(V2 != 332)
```

#### 2) Create a list that contains a country name, the average satisfaction in that country and the observations for Mexico.

```{r tidy = FALSE}

# function to get country code
getCountryCode <- function(name) {
  for(i in 1:NROW(countryCode$country)) {
    if (name == countryCode$country[i]) {
      return (countryCode$V2[i])
    }
  }
}

# create a country name
mexName <- c("Mexico")

# create ave satisfaction
mexCode <- getCountryCode("Mexico")
aveSatisfaction <- mean(wv %>%
                          filter(Country == mexCode) %>%
                          pull(Satisfied))

# observations for Mexico
mexObs <- wv %>%
  filter(wv$Country == mexCode)

# create list containing all the observations for Mexico
mexList = list(countryName = mexName, Satisfaction = aveSatisfaction, observations = mexObs)

# display list
head(mexList)
```

#### 3) Estimate a model predicting *DemocracyBad* (whether someone thinks democracy is bad) as a function of age and save it as OLS.1.  Briefly explain the structure of this object. What is OLS.1[[1]]?

```{r tidy = FALSE}

# lm function
OLS.1 <- lm(
  formula = (wv$DemocracyBad ~ age),
  data= wv
)

# explaination
names(OLS.1)
str(OLS.1)
summary(OLS.1)

# what is OLS.1[[1]]? - coefficients of the model. The two constants represent intercept and slope
OLS.1[[1]]
```
#### 4) Loop thru all countries in the World Values survey to assess the bivariate relationship between opposition to democracy and age. Save coefficient on age and t-statistic on age for each country.

```{r tidy = FALSE}

# extract country code, democracy, age columns from wv
wv2 <- wv %>%
  select(Country, DemocracyBad, age)

# create a hash to store country info (that only exist in wv). key: country code, value: country name
countryInfo <- hash()
for (i in 1:NROW(countryCode)) {
  if(as.character(countryCode$V2[i]) %in% wv$Country) {
    countryInfo[[as.character(countryCode$V2[i])]] <- countryCode$country[i]
  }
}

# create a data frame to store the result
resultDf <- matrix(NA, length(unique(wv$Country)), 4)
colnames(resultDf) <- c("Country Code", "Country", "Coefficient", "T-statistic")

# resultDf <- data.frame(code = character(), country = character(), coeff = double(), tstat = double())

# loop through all countries for democracy vs age using wv2
i = 0
for (code in keys(countryInfo))  {
  i = i + 1
  # create a df for the current country
  curCountryDf <- wv2 %>%
    filter(as.character(Country) == as.character(code)) %>%
    select(DemocracyBad, age)
  # the bivariate relationship between opposition to democracy and age
  if (code == 634) {
    coeff <- NA
    tstat <- NA
    # result = data.frame(code, countryInfo[[code]], NA , NA)
  } else {
    curCountryFit <- lm(
    formula = (DemocracyBad ~ age),
    data = curCountryDf,
    na.action = na.exclude
    )
    coeff <- summary(curCountryFit)$coefficients[2,1]
    tstat <- summary(curCountryFit)$coefficients[2,3]
    # result = data.frame(code, countryInfo[[code]], coeff, tstat)
    # print(result)
  }
  
  # store the current result
  # resultDf <- rbind(resultDf, result)
  resultDf[i, ] <- c(code, countryInfo[[code]], coeff, tstat)
}

# display the result
print(resultDf)
```


#### 5) Create a function that prints a table with response percent for each answer for *DemocracyBad* question for any country

Recall that the answer categories are:

1. Very good
2. Fairly good
3. Fairly bad
4. Very bad  

```{r}

# merge wv and countryCode -> wvJoint
wvJoint <- left_join(wv, countryCode, by = c("Country" = "V2"))

# create a df to store result
resultDf2 <- data.frame(NA)

# calculate proportion and store all the results in resultDf2
for (code in keys(countryInfo)) {
  if (code == 634) {
    next
  }
  # create a df for the current country
  curCountryDf <- wvJoint %>%
    filter(as.character(Country) == as.character(code)) %>%
    select(Country, DemocracyBad)
  # store current result
  curTable <- prop.table(table(wvJoint$DemocracyBad, wvJoint$country), margin = 2) * 100
  # add current result to result
  resultDf2 <- cbind(curTable)
}

# rename the rating names
rownames(resultDf2) = c("Very good", "Fairly good", "Fairly bad", "Very bad")

# function that prints a table with response percent for each answer for *DemocracyBad* question for any country
getResponsePercent <- function(countryName) {
  for (name in colnames(resultDf2)) {
    if (name == countryName) {
      print(resultDf2[, name, drop = FALSE])
    }
  }
}

# test: United States
getResponsePercent("United States")

cat("\n")

# test: China
getResponsePercent("China")

```

#### 6) Create a list with regression results for an OLS model explaining *DemocracyBad* as a function of age, income, education and male.  Be able to print results by country name.

```{r}

countryNames <- unique(wvJoint$country)
countryNames <- subset(countryNames, countryNames != "Qatar")

listRes <- sapply(countryNames, function(name) {
  summary(lm(
    formula = (DemocracyBad ~ age + Income + Education + Male),
    data = wvJoint[name == wvJoint$country, ]
  ))
})

# test: United States
listRes[["United States"]]

```
